---
title: "Hunters Creek Park"
subtitle: "Elma, NY — Trail Conditions Report"
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
# ── Identify this park and load shared infrastructure ──────────────────────
park_id <- "hunters-creek"
source("../../R/setup.R")
source("../../R/parks_config.R")
source("../../R/fetch_weather.R")
source("../../R/fetch_streamflow.R")
source("../../R/fetch_nws.R")
source("../../R/mud_model.R")
source("../../R/snow_model.R")
source("../../R/plots.R")
source("../../R/mudface.R")
source("../../R/snowface.R")
```

```{r fetch-data}
# ── Fetch all data ─────────────────────────────────────────────────────────
weather_raw  <- fetch_weather()
weather_mud  <- compute_mud_level(weather_raw)
stream       <- fetch_streamflow()
nws          <- tryCatch(fetch_nws_forecast(), error = function(e) {
  message("NWS fetch failed: ", e$message); .nws_empty()
})
snow_summary <- current_snow_summary(weather_mud)

# Today's conditions
today_row  <- today_mud(weather_mud)
mud_now    <- today_row$mud_level
mud_name   <- today_row$mud_level_name
depth_cm   <- coalesce(today_row$snow_depth, 0)

# Decide default tab: snow if winter months or snow on ground
is_winter  <- month(Sys.Date()) >= 11 | month(Sys.Date()) <= 4
show_snow  <- is_winter | depth_cm > 2

last_updated <- format(Sys.time(), "%B %d, %Y at %I:%M %p %Z", tz = PARK_TZ)
```

::: {.callout-note appearance="minimal"}
Last updated: `r last_updated`. Data from [Open-Meteo](https://open-meteo.com), [USGS](https://waterdata.usgs.gov), and [NWS](https://www.weather.gov).
:::

::: {.panel-tabset}

## 🥾 Trail Conditions

```{r mud-hero}
#| output: asis

oneliner <- mud_oneliner(today_row, activities = park$activities)

# Large mudface + level badge + summary text
cat('<div class="tp-hero">')
cat('<div class="tp-face">')
cat(as.character(make_mudface(mud_now, size = "180px", label = FALSE)))
cat('</div>')
cat('<div class="tp-hero-text">')
cat(glue('<h2 style="margin-top:0">{coalesce(mud_name, "Unknown")}</h2>'))
cat(glue('<p style="font-size:1.1rem">{oneliner}</p>'))

# Quick stats row
precip_7d <- weather_mud |>
  filter(date >= Sys.Date() - 7, date <= Sys.Date()) |>
  summarise(total = sum(precip, na.rm = TRUE)) |>
  pull(total) |>
  round(1)

temp_today <- coalesce(today_row$temp_max, NA_real_)

cat('<div style="display:flex;gap:1.5rem;flex-wrap:wrap;margin-top:0.75rem;">')
if (!is.na(temp_today)) {
  cat(glue('<div class="tp-stat-box">',
           '<div class="stat-value">{round(temp_today)}°C</div>',
           '<div class="stat-label">Today\'s High</div>',
           '</div>'))
}
cat(glue('<div class="tp-stat-box">',
         '<div class="stat-value">{precip_7d} mm</div>',
         '<div class="stat-label">Precip Last 7 Days</div>',
         '</div>'))
if (!is.na(mud_now)) {
  cat(glue('<div class="tp-stat-box">',
           '<div class="stat-value" style="color:{mud_level_color(mud_now)}">{round(mud_now, 1)}</div>',
           '<div class="stat-label">Mud Score (0–10)</div>',
           '</div>'))
}
cat('</div>')  # stats row
cat('</div>')  # hero text
cat('</div>')  # hero
```

### 14-Day Window + 7-Day Forecast

```{r mud-forecast-chart}
plot_mud_forecast(weather_mud)
```

```{r mud-forecast-strip}
#| output: asis
cat("<p><strong>Forecast mud levels:</strong></p>")
mud_forecast_strip(weather_mud)
```

```{r nws-forecast}
#| output: asis
if (nrow(nws) > 0) {
  cat('<details><summary><strong>NWS 7-Day Forecast</strong></summary>')
  cat('<div style="margin-top:0.75rem;">')
  cat(as.character(nws_forecast_strip(nws)))
  cat('</div></details>')
}
```

---

### Compared to Previous Years

```{r precip-percentile}
#| output: asis

# Cumulative precip so far this calendar year vs. historical
doy_now <- yday(Sys.Date())
ytd_now <- weather_mud |>
  filter(year(date) == year(Sys.Date()), date <= Sys.Date(), source == "history") |>
  summarise(total = sum(precip, na.rm = TRUE)) |>
  pull(total)

hist_ytd <- weather_mud |>
  filter(year(date) < year(Sys.Date()), source == "history") |>
  mutate(doy = yday(date)) |>
  filter(doy <= doy_now) |>
  group_by(yr = year(date)) |>
  summarise(total = sum(precip, na.rm = TRUE), .groups = "drop")

pct_rank <- if (nrow(hist_ytd) > 0)
  round(100 * mean(hist_ytd$total < ytd_now)) else NA_real_

if (!is.na(pct_rank)) {
  direction <- if (pct_rank >= 50) "wetter" else "drier"
  cat(glue('<div class="tp-percentile" style="margin-bottom:1rem;">',
           'Year-to-date precipitation is <strong>{direction} than {pct_rank}%</strong> ',
           'of years since {year(WEATHER_START)}.',
           '</div>'))
}
```

#### Cumulative Precipitation — All Years

```{r precip-spaghetti}
plot_cumulative_precip_ytd(weather_mud)
```

#### Soil Wetness History

```{r soil-moisture}
plot_soil_moisture_history(weather_mud)
```

#### Streamflow — Buffalo River near Elma

```{r streamflow}
plot_streamflow(stream)
```

## ❄️ Snow & Winter

```{r snow-hero}
#| output: asis

snow_one <- snow_oneliner(depth_cm, snow_summary$snow_pct_rank)

cat('<div class="tp-hero">')
cat('<div class="tp-face">')
cat(as.character(make_snowface(depth_cm, size = "180px", label = FALSE)))
cat('</div>')
cat('<div class="tp-hero-text">')
cat(glue('<h2 style="margin-top:0">{snow_level_name(depth_cm)}</h2>'))
cat(glue('<p style="font-size:1.1rem">{snow_one}</p>'))

# Quick stats row
total_snow_this_season <- coalesce(snow_summary$cum_snowfall_cm, 0)
snow_days_this_season  <- if (nrow(snow_summary$current_season) > 0)
  sum(snow_summary$current_season$snowfall > 1, na.rm = TRUE) else 0

cat('<div style="display:flex;gap:1.5rem;flex-wrap:wrap;margin-top:0.75rem;">')
cat(glue('<div class="tp-stat-box">',
         '<div class="stat-value">{round(depth_cm)} cm</div>',
         '<div class="stat-label">Snow Depth Now</div>',
         '</div>'))
cat(glue('<div class="tp-stat-box">',
         '<div class="stat-value">{round(total_snow_this_season)} cm</div>',
         '<div class="stat-label">Season Total</div>',
         '</div>'))
cat(glue('<div class="tp-stat-box">',
         '<div class="stat-value">{snow_days_this_season}</div>',
         '<div class="stat-label">Snow Days</div>',
         '</div>'))
cat('</div>')  # stats row
cat('</div>')  # hero text
cat('</div>')  # hero
```

### 14-Day Window + 7-Day Forecast

```{r snow-forecast-chart}
plot_snow_forecast(weather_mud)
```

```{r snow-forecast-strip}
#| output: asis
cat("<p><strong>Forecast snow conditions:</strong></p>")
snow_forecast_strip(weather_mud)
```

```{r nws-snow-forecast}
#| output: asis
if (nrow(nws) > 0) {
  cat('<details><summary><strong>NWS 7-Day Forecast</strong></summary>')
  cat('<div style="margin-top:0.75rem;">')
  cat(as.character(nws_forecast_strip(nws)))
  cat('</div></details>')
}
```

---

### Compared to Previous Winters

```{r snow-percentile}
#| output: asis
pct <- snow_summary$snow_pct_rank
if (!is.na(pct)) {
  direction <- if (pct >= 50) "snowier" else "less snowy"
  cat(glue('<div class="tp-percentile" style="margin-bottom:1rem;">',
           'This season has been <strong>{direction} than {pct}%</strong> ',
           'of winters since {year(WEATHER_START)}.',
           '</div>'))
}
```

#### Cumulative Snowfall — All Seasons

```{r snowfall-spaghetti}
plot_cumulative_snowfall(weather_mud)
```

#### Peak Snow Depth by Winter

```{r snow-depth-history}
plot_snow_depth_history(weather_mud)
```

:::

---

::: {.callout-tip collapse="true"}
### About the Mud Model

Trail muddiness is estimated using a weighted average of soil moisture in the top 9 cm
(ERA5-Land reanalysis for history, Open-Meteo forecast model for future days).
A logistic sigmoid maps soil wetness (0–0.6 m³/m³) to a 0–10 mud score.
An additional boost is applied when precipitation exceeds 5 mm in the prior 24 hours.
When surface temperature is below 0°C the trail surface is flagged as **Frozen**.

The model has not been calibrated against field observations at this location —
treat mud scores as estimates, especially near the boundaries between named levels.
:::
